---
title: "ev_mab_coverage"
author: "Sadik M"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

This document describes analysis of NGS reads coverage of enset (**Enset ventricosum**) and banana (**Musa** spp) genotypes against genomes of 


````{r load library, echo=FALSE, }
# Load neccessary R packages

library(tidyverse)
library(ggtext)
library(VennDiagram)
library(cowplot)
library(ggVennDiagram)
library(glue)
library(ggtext)
library(ggrepel)

```

# Set input/out directory 

```{r set input_output_dir, echo=FALSE}
# DATA_DIR="/Users/u1866313/allProject_sciptDir/local_scripts/ev_mab_comparative_genomics/data_dir/"
# DATA_DIR="D:/persdoc_dir/allProject_sciptDir/local_scripts/ev_mab_comparative_genomics/data_dir/"
DATA_DIR="Z:/sadik/evmab_dir/data_dir/"

# create output directory 
dir.create(paste0(paste0( DATA_DIR,"../", "output_dir")))
OUTPUT_DIR=paste0(paste0( DATA_DIR,"../", "output_dir"))

```


## Function for reading reads coverage  
```{r read data, echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
reads_coverage <- function(data_dir = "files_directory",
                           ref_genome = "ref_genome",
                           # file_prefix="files_prefix",
                           file_ending_pattern = "files ending pattern including the last separator",
                           strings_to_remove = "strings to remove after the prefix") {
  # lists of reads files
  
  COV_LIST = list.files(path = paste0(DATA_DIR, ref_genome, "/"),
                        pattern = file_ending_pattern) 
  # Coverage of landraces query genotype (landraces or genotypes) against reference genome (ref_genome)
  
  reads_coverage_out = list()
  
  
  # reads data
  for (i in COV_LIST) {
    BASENAME <- basename(i) %>%
      str_remove(paste0(ref_genome, "_")) %>% str_remove(strings_to_remove)
    
    read_cov <- read.delim(paste0(DATA_DIR, ref_genome, "/", i),
                           header = F,
                           sep = '\t') %>%
      rename(
        seq.name = V1,
        start = V2,
        end = V3,
        overlap_count = V4,
        overlap_length = V5,
        total_length_ref = V6,
        fraction_overlap = V7
      ) %>%
      mutate(ref_genome = ref_genome, reads_source = BASENAME)
    
    
    reads_coverage_out[[i]] <- read_cov # add it to your list
    
  }
  
  # Unlist
  
  reads_coverage_out =
    do.call(rbind, reads_coverage_out) %>%
    separate(reads_source,
             sep = "\\.",
             into = c("landrace", "reads_used"))
  
  # remove row.names()
  
  row.names(reads_coverage_out) <- NULL
  rm(read_cov)
  
  # save output as r object
  
  return(reads_coverage_out)
  
}
```


## Reads coverage data: Interspecies reads mapping

### Musa acuminata (MA) genome vs Ensete ventircosum (EV) genotypes 
```{r reads data, echo=FALSE, error=FALSE, warning=FALSE,message=FALSE}

# For Musa acuminata
ref_genome = "musa_acuminata"
file_ending_pattern="cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panEVReads_MAgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


panEVReads_MAgenome_cov <- 
panEVReads_MAgenome_cov %>%
  left_join(
read.delim(paste0(DATA_DIR,"Musa_acuminata.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"Musa_acuminata.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


```

### Musa balbisiana (MB) vs EV genotypes
```{r}
ref_genome = "musa_balbisiana"
file_ending_pattern="cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panEVReads_MBgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


panEVReads_MBgenome_cov <- 
panEVReads_MBgenome_cov %>%
  left_join(
read.delim(paste0(DATA_DIR,"Musa_balbisiana.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"Musa_balbisiana.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 

```

### EV landrace Mazia vs MA and MB genotypes
```{r}
ref_genome = "ev_mazia"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panMABReads_EVMZgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panMABReads_EVMZgenome_cov <-
panMABReads_EVMZgenome_cov %>%
  mutate(landrace = str_remove(landrace,"EV_mazia_")) %>%
  left_join(
read.delim(paste0(DATA_DIR,"EV_mazia.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"EV_mazia.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 

# Remainings


ref_genome = "ev_mazia"
file_ending_pattern=".gene.cov.rem.bed"
strings_to_remove=".gene.cov.rem.bed"


panMABReads_EVMZgenome_cov_rem <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


# combine 
panMABReads_EVMZgenome_cov <-
panMABReads_EVMZgenome_cov %>%
  select(-reads_genome) %>%
  rbind(
panMABReads_EVMZgenome_cov_rem %>%
  separate(seq.name, into = c("seq.name","coords"), sep = ":") %>%
  separate(coords, into = c("start","end"), sep = "-") %>%
  mutate(landrace = str_remove(landrace,"EV_mazia_"),
         gene_repeat = "gene")) 

panMABReads_EVMZgenome_cov <- 
panMABReads_EVMZgenome_cov %>%
  distinct()
  

rm (panMABReads_EVMZgenome_cov_rem)



# Using MA longreads 

ref_genome = "ev_mazia"
file_ending_pattern="_EV_mazia.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMALongReads_EVMZgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>%
    mutate(reads_used = str_remove(reads_used,"_EV_\\w+"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_"))

panMALongReads_EVMZgenome_cov %>%
  head()


# Using MB longreads 

ref_genome = "ev_mazia"
file_ending_pattern="MB.pacbio_EV_mazia.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMBLongReads_EVMZgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>% 
    mutate(reads_used = str_remove(reads_used,"_EV_\\w+"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_"))

panMBLongReads_EVMZgenome_cov %>%
  head()
```

### EV landrace Bedadeti vs MA and MB genotypes

```{r}
ref_genome = "ev_bedadeti"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panMABReads_EVBDgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panMABReads_EVBDgenome_cov <-
panMABReads_EVBDgenome_cov %>%
  mutate(landrace = str_remove(landrace,"EV_bedadeti_")) %>%
    left_join(
read.delim(paste0(DATA_DIR,"EV_bedadeti.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"EV_bedadeti.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


# Remainings


ref_genome = "ev_bedadeti"
file_ending_pattern=".gene.cov.rem.bed"
strings_to_remove=".gene.cov.rem.bed"


panMABReads_EVBDgenome_cov_rem <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


# combine 
panMABReads_EVBDgenome_cov <-
panMABReads_EVBDgenome_cov %>%
  select(-reads_genome) %>%
  rbind(
panMABReads_EVBDgenome_cov_rem %>%
  separate(seq.name, into = c("seq.name","coords"), sep = ":") %>%
  separate(coords, into = c("start","end"), sep = "-") %>%
  mutate(landrace = str_remove(landrace,"EV_bedadeti_"),
         gene_repeat = "gene")) 

panMABReads_EVBDgenome_cov <- 
panMABReads_EVBDgenome_cov %>%
  distinct()
  

rm (panMABReads_EVBDgenome_cov_rem)



# Using MA longreads 

ref_genome = "ev_bedadeti"
file_ending_pattern="_EV_bedadeti.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMALongReads_EVBDgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>%
  mutate(reads_used = str_remove(reads_used,"_EV_\\w+"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_"))
  # mutate(landrace = str_replace(landrace,"MA","MA_longreads"),
  #        landrace = str_replace(landrace,"MS","MS_longreads"),
  #        landrace = str_replace(landrace,"MT","MT_longreads"))

panMALongReads_EVBDgenome_cov %>%
  head()



# Using MB longreads 

ref_genome = "ev_bedadeti"
file_ending_pattern="MB.pacbio_EV_bedadeti.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMBLongReads_EVBDgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>%
  mutate(reads_used = str_remove(reads_used,"_EV_\\w+"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_"))
  # mutate(landrace = str_replace(landrace,"MA","MA_longreads"),
  #        landrace = str_replace(landrace,"MS","MS_longreads"),
  #        landrace = str_replace(landrace,"MT","MT_longreads"))

panMBLongReads_EVBDgenome_cov %>%
  head()
```

## Reads coverage data: Intraspecies reads mapping

### MA vs MA genotypes 
```{r}
ref_genome = "musa_acuminata_genotypes"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panMAReads_MAgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panMAReads_MAgenome_cov <-
panMAReads_MAgenome_cov %>%
  mutate(landrace = str_remove(landrace,"Musa_acuminata_")) %>%
  mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%
    left_join(
read.delim(paste0(DATA_DIR,"Musa_acuminata.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"Musa_acuminata.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


# MA longreads

ref_genome = "musa_acuminata_genotypes"
file_ending_pattern="_Musa_acuminata.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMALongReads_MAgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>%
    mutate(reads_used = str_remove(reads_used,"_Musa_\\w+"),
           reads_used = str_replace(reads_used,"hifi","PacBio_HiFi"),
           reads_used = str_replace(reads_used,"ont","ONT"),
           ref_genome = str_remove(ref_genome,"_genotypes"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_")) 





```

### MB vs MB genotypes 
```{r}
ref_genome = "musa_balbisiana_genotypes"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panMBReads_MBgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panMBReads_MBgenome_cov <-
panMBReads_MBgenome_cov %>%
  mutate(landrace = str_remove(landrace,"Musa_balbisiana_")) %>%
    mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%

  left_join(
read.delim(paste0(DATA_DIR,"Musa_balbisiana.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"Musa_balbisiana.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


# MB longreads

ref_genome = "musa_balbisiana_genotypes"
file_ending_pattern="_Musa_balbisiana.allMapped.gene.cov.bed"
strings_to_remove=".allMapped.gene.cov.bed"


panMBLongReads_MBgenome_cov <-
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                ) %>%
    mutate(reads_used = str_remove(reads_used,"_Musa_\\w+"),
           reads_used = str_replace(reads_used,"hifi","PacBio_HiFi"),
           reads_used = str_replace(reads_used,"pacbio","PacBio_CLR"),
           reads_used = str_replace(reads_used,"ont","ONT"),
           ref_genome = str_remove(ref_genome,"_genotypes"),
         landrace = str_c(landrace,reads_used,"longreads", sep = "_")) 


panMBLongReads_MBgenome_cov %>%
  distinct(landrace)
```


###  EV landrace Bedadeti vs EV genotypes
```{r}
ref_genome = "ev_bedadeti_genotypes"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panEVReads_EVBDgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panEVReads_EVBDgenome_cov <-
panEVReads_EVBDgenome_cov %>%
  mutate(landrace = str_remove(landrace,"EV_bedadeti_")) %>%
  mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%
      left_join(
read.delim(paste0(DATA_DIR,"EV_bedadeti.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"EV_bedadeti.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


ref_genome = "ev_bedadeti_genotypes"
file_ending_pattern=".gene.cov.rem.bed"
strings_to_remove=".gene.cov.rem.bed"


# Remaining
panEVReads_EVBDgenome_cov_rem <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


# combine 
panEVReads_EVBDgenome_cov <-
panEVReads_EVBDgenome_cov %>%
  # select(-reads_genome) %>%
  rbind(
panEVReads_EVBDgenome_cov_rem %>%
  separate(seq.name, into = c("seq.name","coords"), sep = ":") %>%
  separate(coords, into = c("start","end"), sep = "-") %>%
  mutate(landrace = str_remove(landrace,"EV_bedadeti_"),
         gene_repeat = "gene")) 

panEVReads_EVBDgenome_cov <- 
panEVReads_EVBDgenome_cov %>%
      mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%
  distinct()
  

rm (panEVReads_EVBDgenome_cov_rem)

panEVReads_EVBDgenome_cov %>%
  tail()
```

###  EV landrace mazia vs EV genotypes
```{r}
ref_genome = "ev_mazia_genotypes"
file_ending_pattern=".gene_repeats.cov.bed"
strings_to_remove=".gene_repeats.cov.bed"


panEVReads_EVMZgenome_cov <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )

panEVReads_EVMZgenome_cov <-
panEVReads_EVMZgenome_cov %>%
  mutate(landrace = str_remove(landrace,"EV_mazia_")) %>%
  mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%
    left_join(
read.delim(paste0(DATA_DIR,"EV_mazia.gene.bed"), header = F) %>%
  mutate(gene_repeat="gene") %>%
  rbind(
    read.delim(paste0(DATA_DIR,"EV_mazia.gene.complement.bed"), header = F) %>%
  mutate(gene_repeat="repeat")) %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)) 


# Remainings


ref_genome = "ev_mazia_genotypes"
file_ending_pattern=".gene.cov.rem.bed"
strings_to_remove=".gene.cov.rem.bed"


panEVReads_EVMZgenome_cov_rem <- 
reads_coverage(data_dir = DATA_DIR,
               ref_genome = ref_genome,
               file_ending_pattern = file_ending_pattern,
               strings_to_remove = strings_to_remove 
                )


# combine 
panEVReads_EVMZgenome_cov <-
panEVReads_EVMZgenome_cov %>%
  # select(-reads_genome) %>%
  rbind(
panEVReads_EVMZgenome_cov_rem %>%
  separate(seq.name, into = c("seq.name","coords"), sep = ":") %>%
  separate(coords, into = c("start","end"), sep = "-") %>%
  mutate(landrace = str_remove(landrace,"EV_mazia_"),
         gene_repeat = "gene")) 

panEVReads_EVMZgenome_cov <- 
panEVReads_EVMZgenome_cov %>%
    mutate(ref_genome = str_remove(ref_genome,"_genotypes")) %>%
  distinct()
  

rm (panEVReads_EVMZgenome_cov_rem)
panEVReads_EVMZgenome_cov %>%
  tail()
```

## Checking and tidying up data
```{r}

## partitioning reads source 

panMABReads_EVMZgenome_cov <-
  panMABReads_EVMZgenome_cov %>%
    mutate(reads_genome = case_when(landrace == "ERR3412984"|
                                    landrace == "ERR10695529"|
                                    landrace == "ERR10695603"|
                                    landrace == "ERR10695605"|
                                    landrace == "ERR10695608"|
                                    landrace == "SRR6996489"|
                                    landrace == "SRR956987"|
                                    landrace == "SRR957627"|
                                    landrace == "ERR10695529"|
                                    landrace == "ERR10695565"|
                                    landrace == "SRR6147592"|
                                    landrace == "SRR6996488"|
                                    landrace == "SRR8989633"|
                                    landrace == "SRR9734079"|
                                    landrace == "SRR9850640" ~ "B-sub genome banana genotypes",
                                  TRUE ~ "A-sub genome banana genotypes") ) 


# Update with long reads mapping 

panMABReads_EVMZgenome_cov <-
  panMABReads_EVMZgenome_cov %>%
  rbind(
    panMALongReads_EVMZgenome_cov %>%
      rbind(panMBLongReads_EVMZgenome_cov) %>%
      mutate(reads_genome = case_when(str_detect(landrace, "MB_") ~ "B-sub genome banana genotypes",
                                    str_detect(landrace, "MA_") ~ "A-sub genome banana genotypes",
                                    str_detect(landrace, "MS_") ~ "S-sub genome banana genotypes",
                                    str_detect(landrace, "MT_") ~ "T-sub genome banana genotypes" ),
             gene_repeat = "gene")) %>%
  distinct()
    


    
  
panMABReads_EVBDgenome_cov <- 
panMABReads_EVBDgenome_cov %>%
  mutate(reads_genome = case_when(landrace == "ERR3412984"|
                                    landrace == "ERR10695529"|
                                    landrace == "ERR10695603"|
                                    landrace == "ERR10695605"|
                                    landrace == "ERR10695608"|
                                    landrace == "SRR6996489"|
                                    landrace == "SRR956987"|
                                    landrace == "SRR957627"|
                                    landrace == "ERR10695529"|
                                    landrace == "ERR10695565"|
                                    landrace == "SRR6147592"|
                                    landrace == "SRR6996488"|
                                    landrace == "SRR8989633"|
                                    landrace == "SRR9734079"|
                                    landrace == "SRR9850640" ~ "B-sub genome banana genotypes",
                                  TRUE ~ "A-sub genome banana genotypes") )  
  


# Update EV with long reads mapping 

panMABReads_EVBDgenome_cov <-
  panMABReads_EVBDgenome_cov %>%
  rbind(
    panMALongReads_EVBDgenome_cov %>%
            rbind(panMBLongReads_EVBDgenome_cov) %>%
      mutate(reads_genome = case_when(str_detect(landrace, "MB_") ~ "B-sub genome banana genotypes",
                                    str_detect(landrace, "MA_") ~ "A-sub genome banana genotypes",
                                    str_detect(landrace, "MS_") ~ "S-sub genome banana genotypes",
                                    str_detect(landrace, "MT_") ~ "T-sub genome banana genotypes" ),
             gene_repeat = "gene")) %>% 
  distinct()
    



# Update MA with long reads mapping 

panMAReads_MAgenome_cov %>%
  rbind(
panMALongReads_MAgenome_cov %>%
    mutate(gene_repeat = "gene")) %>%
  mutate(reads_genome = case_when(str_detect(landrace, "MA_") ~ "A-sub genome banana genotypes" )) %>% 
  head()

# Update MB with long reads mapping 

panMBReads_MBgenome_cov %>%
  rbind(
panMBLongReads_MBgenome_cov %>%
  mutate(gene_repeat = "gene")) %>%
  mutate(reads_genome = case_when(str_detect(landrace, "MB_") ~ "B-sub genome banana genotypes" )) %>% 
  head()


## combine all them together 
EV_MAB_reads_coverage <-
panEVReads_MAgenome_cov %>%
  mutate(query = "panEV_reads",
         reads_genome = "EV genotypes") %>%
  rbind(
    panEVReads_MBgenome_cov %>%
      mutate(query = "panEV_reads",
             reads_genome = "EV genotypes"),
    
    panMABReads_EVMZgenome_cov %>%
      mutate(
        query = case_when(
          reads_genome == "A-sub genome banana genotypes" |
            reads_genome == "B-sub genome banana genotypes" ~ "panMAB_reads"
        ),
        query = case_when(reads_genome == "S-sub genome banana genotypes" ~ "MS_LongReads"),
        query = case_when(reads_genome == "T-sub genome banana genotypes" ~ "MT_LongReads")
      ),
    
    panMABReads_EVBDgenome_cov %>%
      mutate(
        query = case_when(
          reads_genome == "A-sub genome banana genotypes" |
            reads_genome == "B-sub genome banana genotypes" ~ "panMAB_reads"
        ),
        query = case_when(reads_genome == "S-sub genome banana genotypes" ~ "MS_LongReads"),
        query = case_when(reads_genome == "T-sub genome banana genotypes" ~ "MT_LongReads") ),
    # mutate(query = "panMAB_reads"),
    
    panMAReads_MAgenome_cov %>%
      rbind(panMALongReads_MAgenome_cov %>%
              mutate(gene_repeat = "gene")) %>%
      mutate(reads_genome = "A-sub genome banana genotypes",
             query = "panMA_reads"),
    
    panMBReads_MBgenome_cov %>%
      rbind(panMBLongReads_MBgenome_cov %>%
              mutate(gene_repeat = "gene")) %>%
      mutate(reads_genome = "B-sub genome banana genotypes",
             query = "panMB_reads"),
    
    panEVReads_EVBDgenome_cov %>%
      mutate(query = "panEV_reads",
             reads_genome = "EV genotypes"),
    
    panEVReads_EVMZgenome_cov %>%
      mutate(query = "panEV_reads",
             reads_genome = "EV genotypes")
  ) %>%
  filter(gene_repeat=="gene")
  # group_by(seq.name,
  #          start,
  #          end,
  #          ref_genome,
  #          reads_used,
  #          gene_repeat,
  #          query,
  #          reads_genome) %>%
  # summarise(
  #   overlap_count = mean(round(overlap_count, 1)),
  #   overlap_length = mean(round(overlap_length, 1)),
  #   total_length_ref = mean(round(total_length_ref, 1)),
  #   fraction_overlap = mean(round(fraction_overlap, 1))
  # ) %>%
  # ungroup() 

EV_MAB_reads_coverage %>%
  filter(ref_genome=="musa_balbisiana") %>%
  head()
```

## Get regions in Enset genome syntenic to MA and MB that were inferrred from lastz and Mummer alignment 
```{r}
# DATA_DIR

EV_MAB_lastz_nucmer_coding_seq <- 
read.delim(paste0(DATA_DIR,"bedadeti_MA.MA.lastz_nucmer.intersect.bed"), header=F) %>%
  mutate(ref_genome = "musa_acuminata",
         query_genome = "ev_bedadeti",
         syn_cov = round(V7/(V3-V2+1),2)) %>%
  bind_rows(
    
    read.delim(paste0(DATA_DIR,"bedadeti_MB.MB.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "musa_balbisiana",
             query_genome = "ev_bedadeti",
             syn_cov = round(V7/(V3-V2+1),2)),
    
    read.delim(paste0(DATA_DIR,"mazia_MA.MA.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "musa_acuminata",
             query_genome = "ev_mazia",
             syn_cov = round(V7/(V3-V2+1),2)) ,
    read.delim(paste0(DATA_DIR,"mazia_MB.MB.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "musa_balbisiana",
             query_genome = "ev_mazia",
             syn_cov = round(V7/(V3-V2+1),2)) ,
    read.delim(paste0(DATA_DIR,"mazia_MA.mazia.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "ev_mazia",
             query_genome = "musa_acuminata",
             syn_cov = round(V7/(V3-V2+1),2)) ,
    read.delim(paste0(DATA_DIR,"mazia_MB.mazia.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "ev_mazia",
             query_genome = "musa_balbisiana",
             syn_cov = round(V7/(V3-V2+1),2)) ,
    read.delim(paste0(DATA_DIR,"bedadeti_MA.bedadeti.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "ev_bedadeti",
             query_genome = "musa_acuminata",
             syn_cov = round(V7/(V3-V2+1),2)),
    
    read.delim(paste0(DATA_DIR,"bedadeti_MB.bedadeti.lastz_nucmer.intersect.bed"), header=F) %>%
      mutate(ref_genome = "ev_bedadeti",
             query_genome = "musa_balbisiana",
             syn_cov = round(V7/(V3-V2+1),2))) 


```

```{r}

# EV_MAB_reads_coverage_no_lastz_nucmer_alignment <-
# EV_MAB_lastz_nucmer_coding_seq %>%
#   # filter(cov < 0 ) %>%
#   filter(V7 < 1) %>%
#   dplyr::rename(seq.name= V1,
#                 start = V2,
#                 end = V3) %>% 
#   mutate(source="non_syntenic") %>%
#   left_join(
#     EV_MAB_reads_coverage) %>%
#   mutate(reads_genome = str_replace_na(reads_genome,"NA")) %>%
#   as.data.frame() 


# updated

EV_MAB_reads_coverage_no_lastz_nucmer_alignment <-
  
  # Select inter-species reads mapping coverage and join it with nucmer+lastz coding sequence alignment
  EV_MAB_reads_coverage %>%
  as.data.frame() %>%
  mutate(reads_genome = str_replace_na(reads_genome, "NA"),
             start = as.numeric(start),
             end = as.numeric(end)) %>%  mutate(
    mapping_reads_source = case_when(
      ref_genome == "ev_mazia" &
        reads_genome == "EV genotypes" ~ "self_reads",
      ref_genome == "ev_bedadeti" &
        reads_genome == "EV genotypes" ~ "self_reads",
      ref_genome == "musa_acuminata" &
        reads_genome == "A-sub genome banana genotypes" ~ "self_reads",
      ref_genome == "musa_balbisiana" &
        reads_genome == "B-sub genome banana genotypes" ~ "self_reads",
      TRUE ~ "cross_species_reads"
    )
  ) %>%
  filter(mapping_reads_source == "cross_species_reads") %>%
  left_join(
    EV_MAB_lastz_nucmer_coding_seq %>%
      # filter(cov < 0 ) %>%
      # filter(V7 < 1) %>%
      dplyr::rename(
        seq.name = V1,
        start = V2,
        end = V3
      ) %>%
      mutate(synteny = case_when(V6 <= 0 ~ "non_syntenic",
                                 TRUE ~ "syntenic"))
  ) %>%
  # select(-mapping_reads_source) %>%
  # str()
  # concatenate reads intera-species mapping coverage data
  bind_rows(
    EV_MAB_reads_coverage %>%
      as.data.frame() %>%
      mutate(reads_genome = str_replace_na(reads_genome, "NA"),
             start = as.numeric(start),
             end = as.numeric(end)) %>%
      mutate(
        mapping_reads_source = case_when(
          ref_genome == "ev_mazia" &
            reads_genome == "EV genotypes" ~ "self_reads",
          ref_genome == "ev_bedadeti" &
            reads_genome == "EV genotypes" ~ "self_reads",
          ref_genome == "musa_acuminata" &
            reads_genome == "A-sub genome banana genotypes" ~ "self_reads",
          ref_genome == "musa_balbisiana" &
            reads_genome == "B-sub genome banana genotypes" ~ "self_reads",
          TRUE ~ "cross_species_reads"
        )
      ) %>%
      filter(mapping_reads_source != "cross_species_reads") %>%
      # select(-mapping_reads_source) %>%
      mutate(
        V4 = "NA",
        V5 = "NA",
        V6 = "NA",
        V7 = "NA",
        query_genome = "NA",
        syn_cov = "NA",
        synteny = "self_reads",
        V5 = as.integer(V5),
        V6 = as.integer(V7),
        V7 = as.integer(V7),
        syn_cov = as.numeric(syn_cov)
        
      )
  )

EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
  filter(query_genome=="NA")
  # distinct(reads_genome)
  # filter(synteny=="non_syntenic") %>%
  # select(seq.name,start,end) %>%
  # mutate(syn_cov = str_replace_na(syn_cov,"NA")) %>%
  # filter(syn_cov=="NA") %>%
  # filter(reads_used == "allMapped")
  # tail()
    filter(end=="125076") 



```

```{r}
# save.image(paste0(DATA_DIR,"panEV_MAB_reads_coverage.05.RData"))
# load(paste0(DATA_DIR,"panEV_MAB_reads_coverage.05.RData"))

# save.image(paste0(DATA_DIR,"panEV_MAB_reads_coverage.06.RData"))
# load(paste0(DATA_DIR,"panEV_MAB_reads_coverage.06.RData"))

save.image(paste0(DATA_DIR,"panEV_MAB_reads_coverage.07.RData"))
save.image(paste0(DATA_DIR,"panEV_MAB_reads_coverage_filtered.08.RData"))

```

