---
title: "reads_coverage_go_terms"
author: "Sadik M"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load GO-terms annotation of Enset and Musa spp proteins 
```{r}

load("~/Downloads/OrthoFinder.EV.AED25.MAB.RData")
load("~/Downloads/OrthoFinder.EV.AED25.MAB.monocot.RData")

save(GO_terms_EV_AED25_MAB_v2,file = "GO_terms_EV_AED25_MAB_v2.RData")
save(GO_terms_EV_AED25_MAB_desc.v1,file = "GO_terms_EV_AED25_MAB_desc.v1.RData")


```

## Get gene IDs from annotation files and link with reads coverage across geneic regions and GO-term annotation
```{r}
# library(rtracklayer)

EV_MAB_genes <- list()


for (EACH_GFF_ANNOTATION in list.files(path = DATA_DIR, pattern = "gff3")){
  
  BASENAME <- basename(EACH_GFF_ANNOTATION) %>% str_remove(".gff3") %>% str_remove("EV_")

  read_gff <-
    rtracklayer::readGFF(paste0(DATA_DIR,EACH_GFF_ANNOTATION)) %>%
  as.data.frame() %>%
  dplyr::filter(type =='gene') %>%
  select(seqid,start,end,Name) %>%
    mutate(genome = BASENAME)

    EV_MAB_genes [[EACH_GFF_ANNOTATION]] <- read_gff # add it to your list

  
}

rm(read_gff)

EV_MAB_genes =
    do.call(rbind, EV_MAB_genes) %>%
  mutate(genome = str_remove(genome,".maker.annotation"))


# remove rownames 
rownames(EV_MAB_genes) <- NULL


# combine gene ID and reads coverage 
EV_MAB_reads_coverage_geneID <-
EV_MAB_genes %>% 
mutate(start = start -1) %>%
  left_join(
    EV_MAB_reads_coverage %>%
      mutate(
        
        reads_genome = str_replace_na(reads_genome, "NA"),
             start = as.numeric(start),
             end = as.numeric(end),
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name)) 
  # filter(start==185298)

```


## add IntreProScan GO-terms annotation 
```{r}

EV_MAB_reads_coverage_geneID_GO_terms <-
  EV_MAB_reads_coverage_geneID %>%
  # rename gene ID of Musa acuminata (MA) and Musa balbisiana (MB) to match it with gene ID used in GO-terms analysis
  # Simple add a suffix of ".1" to each of MA and MB genes
  mutate(Name = case_when(genome == "Musa_acuminata" |
                            genome == "Musa_balbisiana" ~ str_c(Name,".1"),
                          TRUE ~ Name)) %>%
  dplyr::rename(genes=Name) %>%
  # Join GO-terms annotation
  left_join(
    GO_terms_EV_AED25_MAB_v2 %>% 
      mutate(genomes = str_remove(genomes,".AED25"),
             genes = str_remove(genes,"-\\w+")) %>%
      dplyr::rename(genome = genomes) %>%
      distinct()) %>%
  # filter(genome=="Musa_acuminata") %>% 
  mutate(GO_terms = str_replace_na(GO_terms,"NA")) %>%
  dplyr::rename(gene_id = genes)
  # head() %>%
  # filter(GO_terms != "NA")
```


## Filter EV and MAB specific genes 
### EV mazia specific genes 
```{r}
EV_MAB_reads_coverage_geneID_GO_terms %>%
  filter(ref_genome=="ev_mazia") %>%
  filter(fraction_overlap < 0.25 ) %>%
  head()
```

## update GO-terms annotation for non-syntenic genes

```{r}

# combine gene ID and reads coverage 

# EV_MAB_reads_coverage_geneID_synteny <-
# EV_MAB_genes %>% 
# mutate(start = start -1) %>%
#   left_join(
#     EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
#   # filter(synteny=="non_syntenic") %>%
#       mutate(
#         genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
#                        ref_genome=="musa_acuminata" ~ "Musa_acuminata",
#                        ref_genome=="ev_mazia" ~ "mazia",
#                        ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
#       dplyr::rename(seqid=seq.name)) %>%
#       mutate(V4 = str_replace_na(V4,'NA')) 

EV_MAB_reads_coverage_geneID_synteny <-
    EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
  # filter(synteny=="non_syntenic") %>%
      mutate(
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name) %>%
  left_join(
    EV_MAB_genes %>% 
      mutate(start = start -1)) %>%
      mutate(V4 = str_replace_na(V4,'NA')) 
    


EV_MAB_reads_coverage_geneID_GO_terms_synteny <-
  EV_MAB_reads_coverage_geneID_synteny %>%
  # rename gene ID of Musa acuminata (MA) and Musa balbisiana (MB) to match it with gene ID used in GO-terms analysis
  # Simple add a suffix of ".1" to each of MA and MB genes
  mutate(Name = case_when(genome == "Musa_acuminata" |
                            genome == "Musa_balbisiana" ~ str_c(Name,".1"),
                          TRUE ~ Name)) %>%
  dplyr::rename(genes=Name) %>%
  # Join GO-terms annotation
  left_join(
    GO_terms_EV_AED25_MAB_v2 %>% 
      mutate(genomes = str_remove(genomes,".AED25"),
             genes = str_remove(genes,"-\\w+")) %>%
      dplyr::rename(genome = genomes) %>%
      distinct()) %>%
  mutate(GO_terms = str_replace_na(GO_terms,"NA")) %>%
  dplyr::rename(gene_id = genes) %>%
  mutate (V7 = str_replace_na(V7,"NA")) %>%
  distinct()
  # head() %>%
  # filter(GO_terms != "NA")



        
      EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
                # filter(end=="125076") 
      mutate(
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name) %>%
                 filter(end=="125076") %>%

  left_join(
    EV_MAB_genes %>% 
      mutate(start = start -1)) %>%
      mutate(V4 = str_replace_na(V4,'NA')) 
      
      
       EV_MAB_genes %>% 
      mutate(start = start -1) %>%
         filter(end=="125076") 

     EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
       head()
```


```{r}

# update data structure 

EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 <-
EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>% 
  as.data.frame() %>%
  filter(gene_repeat=="gene" ) %>% 
  # filter(reads_used =="allMapped") %>%
  select( seq.name, start,  end, ref_genome, reads_used, reads_genome, gene_repeat, query, group, fraction_overlap, syn_cov) %>%
  mutate(syn_cov = str_replace_na(syn_cov,"NA")) %>%
  filter(fraction_overlap >= 0.456) %>%
  mutate(frac_ovrerlap_syn = fraction_overlap) %>%
  distinct() %>%
rbind(
    
  EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>% 
  as.data.frame() %>%
  filter(gene_repeat=="gene") %>% 
  # filter(reads_used =="allMapped") %>%
  select(seq.name, start,  end, ref_genome, reads_used,reads_genome, gene_repeat, query, group, fraction_overlap, syn_cov) %>%
  mutate(syn_cov = str_replace_na(syn_cov,"NA"),
         syn_cov = as.numeric(syn_cov)) %>%
  filter(fraction_overlap < 0.245, syn_cov >=0.245) %>%
    mutate(frac_ovrerlap_syn = syn_cov),

  
    EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>% 
      as.data.frame() %>%
      filter(gene_repeat=="gene") %>% 
      # filter(reads_used =="allMapped") %>%
      select(seq.name, start,  end, ref_genome, reads_used, reads_genome, gene_repeat, query, group, fraction_overlap, syn_cov) %>%
      mutate(syn_cov = str_replace_na(syn_cov,"NA"),
             syn_cov = as.numeric(syn_cov)) %>%
      filter(fraction_overlap < 0.245, syn_cov <0.245) %>%
      mutate(frac_ovrerlap_syn = fraction_overlap) 
     ) %>%
  filter(gene_repeat == "gene") %>%
  select(-gene_repeat) %>%
  distinct()
  # head()


# EV_MAB_reads_coverage %>% distinct(group)  %>% dput()

EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 <- 
EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 %>%
  mutate(
    query = case_when(str_detect(reads_genome,"A-sub") ~ "panMA_reads",
                      str_detect(reads_genome,"B-sub") ~ "panMB_reads",
                      str_detect(reads_genome,"S-sub") ~ "MS_LongReads",
                      str_detect(reads_genome,"T-sub") ~ "MT_LongReads",
                      TRUE ~ query))


EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 <-
  EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 %>%
  mutate(
    group = case_when(
      str_detect(ref_genome, "musa_acuminata") &
        str_detect(query, "panEV_reads") ~ "*E. ventricosum* NGS reads vs <br> MA genome",
      str_detect(ref_genome, "musa_acuminata") &
        str_detect(query, "panMA_reads") ~ "*Musa spp.* A-genome NGS reads vs <br> MA genome",
      str_detect(ref_genome, "musa_balbisiana") &
        str_detect(query, "panEV_reads") ~ "*E. ventricosum* NGS reads vs <br> MB genome",
      str_detect(ref_genome, "musa_balbisiana") &
        str_detect(query, "panMB_reads") ~ "*Musa spp.* B-genome NGS vs <br> MB genome",
      str_detect(ref_genome, "ev_bedadeti") &
        str_detect(query, "panMA_reads") ~ "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Bedadeti) genome",
      str_detect(ref_genome, "ev_bedadeti") &
        str_detect(query, "panMB_reads") ~ "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Bedadeti) genome",
      str_detect(ref_genome, "ev_bedadeti") &
        str_detect(query, "panEV_reads") ~ "*E. ventricosum* NGS reads vs <br> EV (Bedadeti) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "panMA_reads") ~ "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Mazia) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "panMB_reads") ~ "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Mazia) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "panEV_reads") ~ "*E. ventricosum* NGS reads vs <br> EV (Mazia) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "MS_LongReads") ~ "*MS long reads vs <br> EV (Mazia) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "MT_LongReads") ~ "*MT long reads vs <br> EV (Mazia) genome",
      str_detect(ref_genome, "ev_bedadeti") &
        str_detect(query, "MS_LongReads") ~ "*MS long reads vs <br> EV (Bedadeti) genome",
      str_detect(ref_genome, "ev_bedadeti") &
        str_detect(query, "MT_LongReads") ~ "*MT long reads vs <br> EV (Bedadeti) genome",
      str_detect(ref_genome, "ev_mazia") &
        str_detect(query, "MT_LongReads") ~ "*MT long reads vs <br> EV (Bedadeti) genome"
    )
    
  )
  



EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3$group = factor(EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3$group,
                                     levels = c( 
                                                "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Mazia) genome", 
                                                "*Musa spp.* A- and B-genome NGS reads vs <br> EV (Bedadeti) genome", 
                                                "*E. ventricosum* NGS reads vs <br> MA genome", 
                                                "*E. ventricosum* NGS reads vs <br> MB genome",
                                                "*Musa spp.* A-genome NGS reads vs <br> MA genome", 
                                                "*Musa spp.* B-genome NGS vs <br> MB genome", 
                                                "*E. ventricosum* NGS reads vs <br> EV (Mazia) genome",
                                                "*E. ventricosum* NGS reads vs <br> EV (Bedadeti) genome",
                                                "*MS long reads vs <br> EV (Mazia) genome",
                                                "*MT long reads vs <br> EV (Mazia) genome",
                                                "*MS long reads vs <br> EV (Bedadeti) genome",
                                                "*MT long reads vs <br> EV (Bedadeti) genome"))

EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3
  head()
# EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 <-
# EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 %>% 
#     as.data.frame() %>%
#   filter(
#     group !="*Musa spp.* A-genome NGS reads vs <br> MA genome",
#     group !="*Musa spp.* B-genome NGS vs <br> MB genome",
#     group !="*E. ventricosum* NGS reads vs <br> MA genome",
#     group !="*E. ventricosum* NGS reads vs <br> MB genome"
#     # !str_detect(group,"long reads")
#   ) %>% 
#   distinct() %>%
#   filter(reads_used !="qfiltered") %>%
#     group_by( seq.name, start,  end, ref_genome, group) %>%
#     # group_by( seq.name, start,  end, ref_genome, reads_used, gene_repeat, query, group) %>%
#   summarise(frac_ovrerlap_syn = max(frac_ovrerlap_syn)) %>%
#   ungroup()
```


## update GO-terms annotation for non-syntenic genes

```{r}

# combine gene ID and reads coverage 


EV_MAB_reads_coverage_geneID_synteny.v1 <-
    EV_MAB_reads_coverage_no_lastz_nucmer_alignment.v3 %>%
  # filter(synteny=="non_syntenic") %>%
      mutate(
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name) %>%
  left_join(
    EV_MAB_genes %>% 
      mutate(start = start -1)) 


EV_MAB_reads_coverage_geneID_GO_terms_synteny.v1 <-
  EV_MAB_reads_coverage_geneID_synteny.v1 %>%
  # rename gene ID of Musa acuminata (MA) and Musa balbisiana (MB) to match it with gene ID used in GO-terms analysis
  # Simple add a suffix of ".1" to each of MA and MB genes
  mutate(Name = case_when(genome == "Musa_acuminata" |
                            genome == "Musa_balbisiana" ~ str_c(Name,".1"),
                          TRUE ~ Name)) %>%
  dplyr::rename(genes=Name) %>%
  # Join GO-terms annotation
  left_join(
    GO_terms_EV_AED25_MAB_v2 %>% 
      mutate(genomes = str_remove(genomes,".AED25"),
             genes = str_remove(genes,"-\\w+")) %>%
      dplyr::rename(genome = genomes) %>%
      distinct()) %>%
  mutate(GO_terms = str_replace_na(GO_terms,"NA")) %>%
  dplyr::rename(gene_id = genes) %>%
  # mutate (V7 = str_replace_na(V7,"NA")) %>%
  distinct()
  # head() %>%
  # filter(GO_terms != "NA")



        
  #     EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
  #               # filter(end=="125076") 
  #     mutate(
  #       genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
  #                      ref_genome=="musa_acuminata" ~ "Musa_acuminata",
  #                      ref_genome=="ev_mazia" ~ "mazia",
  #                      ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
  #     dplyr::rename(seqid=seq.name) %>%
  #                filter(end=="125076") %>%
  # 
  # left_join(
  #   EV_MAB_genes %>% 
  #     mutate(start = start -1)) %>%
  #     mutate(V4 = str_replace_na(V4,'NA')) 
  #     
  #     
  #      EV_MAB_genes %>% 
  #     mutate(start = start -1) %>%
  #        filter(end=="125076") 

     EV_MAB_reads_coverage_geneID_GO_terms_synteny.v1 %>%
       filter(GO_terms !="NA")
       tail()
```

