---
title: "reads_coverage_go_terms"
author: "Sadik M"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Set out put directory 
```{r}
# cwd=getwd()
cwd="Z:/sadik/evmab_dir/data_dir/EV_MAB_compgeno/EV_MAB_compgeno/"

dir.create(paste0(cwd,"result_dir"))
result_dir=paste0(cwd,"result_dir/")
```

# EV mazia CDS < 25 panMA reads coverage 
## CDS regions that showed less than 0.25% NGS/TGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping from MA)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_mazia") %>%
  filter(query == "panMA_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct(seq.name,start,end) %>% 
  nrow()

```


## Assess support of reads-mapping based EV specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_mazia") %>%
  filter(query == "panEV_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
  nrow()
  write.table(paste0(result_dir,"filtered_ev_mazia_cds_less_25_per_panma_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```


# EV mazia CDS < 25 panMB reads coverage 
## CDS regions that showed less than 0.25% NGS/TGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping from MB)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_mazia") %>%
  filter(query == "panMB_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct() %>%
  nrow()

```


## Assess support of reads-mapping based EV mazia specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_mazia") %>%
  filter(query == "panEV_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
  # nrow()
  write.table(paste0(result_dir,"filtered_ev_mazia_cds_less_25_per_panmb_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```


## test
```{r}
EV_MAB_reads_coverage %>%
  distinct(query)
```

# EV bedadeti CDS < 25 panMA reads coverage 
## CDS regions that showed less than 0.25% NGS/TGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping from MA)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_bedadeti") %>%
  filter(query == "panMA_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct(seq.name,start,end) %>%
  nrow()

```


## Assess support of reads-mapping based EV specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_bedadeti") %>%
  filter(query == "panEV_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
  nrow()
  write.table(paste0(result_dir,"filtered_ev_bedadeti_cds_less_25_per_panma_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```


# EV bedadeti CDS < 25 panMB reads coverage 
## CDS regions that showed less than 0.25% NGS/TGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping from MB)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_bedadeti") %>%
  filter(query == "panMB_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct(seq.name,start,end) %>%
  nrow()

```


## Assess support of reads-mapping based EV mazia specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "ev_bedadeti") %>%
  filter(query == "panEV_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
   nrow()
  write.table(paste0(result_dir,"filtered_ev_bedadeti_cds_less_25_per_panmb_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```

# MA CDS < 25% panEV reads coverage 
## CDS regions that showed less than 0.25% NGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "musa_acuminata") %>%
  filter(query == "panEV_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct(seq.name,start,end) %>%
  nrow()

```
## Assess support of reads-mapping based EV specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "musa_acuminata") %>%
  filter(query == "panMA_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
  # nrow()
  write.table(paste0(result_dir,"filtered_MA_cds_less_25_per_panev_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```


#MB CDS < 25% panEV reads coverage 
## CDS regions that showed less than 0.25% NGS reads mapping coverage (proportion bases in CDS that showed NGS/TGS reads mapping)

```{r}
# ev_genomes_list=c("ev_mazia","ev_bedadeti")

species_specific_by_reads_mapping_cov <-
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "musa_balbisiana") %>%
  filter(query == "panEV_reads") %>%

  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame() %>%
  mutate(start = str_replace_na(start, "NA")) %>%
  filter(start != "NA") %>%
  filter(fraction_overlap < 0.25) %>%
  distinct()

species_specific_by_reads_mapping_cov %>% 
  # select(-fraction_overlap) %>%
  # filter(fraction_overlap == 0) %>%
  distinct(seq.name,start,end) %>%
  nrow()

```


## Assess support of reads-mapping based EV specific CDS against EV reads 
```{r}
species_specific_by_reads_mapping_cov %>% 
  left_join(
EV_MAB_reads_coverage %>% 
  filter(reads_used != "qfiltered" ) %>%
  filter(reads_genome != "S-sub genome banana genotypes",
         reads_genome != "T-sub genome banana genotypes") %>%
  # distinct(query) 
  
  mutate(fraction_overlap = round(fraction_overlap,2)) %>%
  filter(ref_genome == "musa_balbisiana") %>%
  filter(query == "panMB_reads") %>%
  select(seq.name,start,end,fraction_overlap) %>%
  group_by(seq.name,start,end) %>%
  summarise(fraction_overlap_panev = max(fraction_overlap)) %>%
  ungroup() %>%
  as.data.frame()) %>%
  filter(fraction_overlap_panev >= 0.25) %>%
  distinct(seq.name,start,end) %>% 
  nrow()
  write.table(paste0(result_dir,"filtered_MB_cds_less_25_per_panev_cov.bed"),
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t"
              )
```


## Inferred syntenic regions from lastz and numcer CDS alignment
```{r}

EV_MAB_lastz_nucmer_coding_seq %>% 
  rename(seq.name = V1,
         start = V2,
         end = V3) %>%
  filter(query_genome == "ev_bedadeti" |
         query_genome == "ev_mazia") %>%
  # distinct(ref_genome) %>%
  filter(ref_genome  == "musa_acuminata") %>%
  # select(-ref_genome,-query_genome)
  tail()
  
```


## CDS regions with less 0.25% reads mapping coverage but showed inter-species CDS alignment across >=25% of the CDS length

```{r}

# species_specific_by_reads_mapping_cov_but_showed_syntenic_alignment
species_specific_by_reads_mapping_cov %>% 
  mutate(start = as.integer(start),
         end = as.integer(end)) %>%
  
  left_join(
    EV_MAB_lastz_nucmer_coding_seq %>%
  rename(seq.name = V1,
         start = V2,
         end = V3) %>%
  filter(query_genome == "musa_acuminata" |
         query_genome == "musa_balbisiana") %>%
  # distinct(ref_genome) %>%
  filter(ref_genome  == "ev_bedadeti")) %>%
  select(seq.name,start,end,syn_cov) %>%
  group_by(seq.name,start,end) %>%
  summarise(syn_cov = max(syn_cov)) %>%
  filter(syn_cov >= 0.25) %>%
  ungroup() %>%
  as.data.frame() %>%
  distinct() %>%
  nrow()
```
## write to disk: CDS regions with less 0.25% reads mapping coverage but showed inter-species CDS alignment across >=25% of the CDS length

```{r}



musa_acuminata				
musa_balbisiana				
ev_mazia				
ev_bedadeti

ev_bedadeti				
ev_mazia				
musa_acuminata				
musa_balbisiana

# ref_genome="musa_acuminata"
# query_genome="ev_bedadeti"
species_specific_by_reads_mapping_cov %>% 
  filter(ref_genome == "musa_acuminata") %>%
  filter(query == "panEV_reads") %>%
  group_by(seq.name,start,end,ref_genome) %>%
  summarise(fraction_overlap = max(fraction_overlap)) %>%
  ungroup() %>%
  select(seq.name,start,end,fraction_overlap,ref_genome) %>%
  distinct() 
  nrow()

  head()
  mutate(start = as.integer(start),
         end = as.integer(end)) %>%
  left_join(
    EV_MAB_lastz_nucmer_coding_seq %>%
  rename(seq.name = V1,
         start = V2,
         end = V3)
  ) %>% 
  # filter(fraction_overlap < 0.25)
  select(seq.name,start,end,syn_cov,query_genome,ref_genome,fraction_overlap) %>%
  mutate(fraction_overlap = str_replace_na(fraction_overlap, "NA")) %>%
  filter(fraction_overlap != "NA") %>%
  filter(syn_cov >= 0.25) %>%
  # distinct() %>%
  filter(ref_genome == "musa_acuminata") %>%
  filter(query_genome == "ev_bedadeti") %>%
  # distinct(ref_genome)
  group_by(seq.name,start,end) %>%
  summarise(syn_cov = max(syn_cov)) %>%
  ungroup() %>%
  select(seq.name,start,end,syn_cov) %>%
  distinct() %>%
  nrow()
  write.table(paste0(result_dir, "ev_bedadeti", ".", "musa_acuminata", ".lackReadsSupport_showedCDSalign.bed"), 
              col.names = F,
              row.names = F,
              quote = F,
              sep = "\t")


```


```{r}
# combine gene ID and reads coverage 

# EV_MAB_reads_coverage_geneID_synteny <-
# EV_MAB_genes %>% 
# mutate(start = start -1) %>%
#   left_join(
#     EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
#   # filter(synteny=="non_syntenic") %>%
#       mutate(
#         genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
#                        ref_genome=="musa_acuminata" ~ "Musa_acuminata",
#                        ref_genome=="ev_mazia" ~ "mazia",
#                        ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
#       dplyr::rename(seqid=seq.name)) %>%
#       mutate(V4 = str_replace_na(V4,'NA')) 

EV_MAB_reads_coverage_geneID_synteny <-
    EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
  # filter(synteny=="non_syntenic") %>%
      mutate(
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name) %>%
  left_join(
    EV_MAB_genes %>% 
      mutate(start = start -1)) %>%
      mutate(V4 = str_replace_na(V4,'NA')) 
    


EV_MAB_reads_coverage_geneID_GO_terms_synteny <-
  EV_MAB_reads_coverage_geneID_synteny %>%
  # rename gene ID of Musa acuminata (MA) and Musa balbisiana (MB) to match it with gene ID used in GO-terms analysis
  # Simple add a suffix of ".1" to each of MA and MB genes
  mutate(Name = case_when(genome == "Musa_acuminata" |
                            genome == "Musa_balbisiana" ~ str_c(Name,".1"),
                          TRUE ~ Name)) %>%
  dplyr::rename(genes=Name) %>%
  # Join GO-terms annotation
  left_join(
    GO_terms_EV_AED25_MAB_v2 %>% 
      mutate(genomes = str_remove(genomes,".AED25"),
             genes = str_remove(genes,"-\\w+")) %>%
      dplyr::rename(genome = genomes) %>%
      distinct()) %>%
  mutate(GO_terms = str_replace_na(GO_terms,"NA")) %>%
  dplyr::rename(gene_id = genes) %>%
  mutate (V7 = str_replace_na(V7,"NA")) %>%
  distinct()
  # head() %>%
  # filter(GO_terms != "NA")



        
      EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
                # filter(end=="125076") 
      mutate(
        genome = case_when(ref_genome=="musa_balbisiana" ~ "Musa_balbisiana",
                       ref_genome=="musa_acuminata" ~ "Musa_acuminata",
                       ref_genome=="ev_mazia" ~ "mazia",
                       ref_genome=="ev_bedadeti" ~ "bedadeti")) %>%
      dplyr::rename(seqid=seq.name) %>%
                 filter(end=="125076") %>%

  left_join(
    EV_MAB_genes %>% 
      mutate(start = start -1)) %>%
      mutate(V4 = str_replace_na(V4,'NA')) 
      
      
       EV_MAB_genes %>% 
      mutate(start = start -1) %>%
         filter(end=="125076") 

     EV_MAB_reads_coverage_no_lastz_nucmer_alignment %>%
       head()
```



